Understanding the Socio-Economic Determinants of Voter Behavior: An Analysis of the 2020 U.S. Presidential Election

### 2024-03-12 paper/us_election_2020.qmd
model implication

---
title: "Understanding the Socio-Economic Determinants of Voter Behavior: An Analysis of the 2020 U.S. Presidential Election"
author: 
  - Hannah Yu
  - Mary Cheng
  - Yimiao Yuan
thanks: "Code and data are available at: https://github.com/hannahyu07/US-Election"
date: today
date-format: long
abstract: "This paper delves into the dynamics of voter turnout of the 2020 United States presidential election. Utilizing dataset from the Cooperate Election Study (CES2020), we employ logistic regression analysis to explore the influence of the voter's race, region, and employment on their vote for presidential candidate. Our findings reveal distinct voting patterns among different demographic groups, with minorities showing strong support for Biden, the South predominantly favored Trump, and individuals reliant on unemployment benefits preferred Biden. These insights stress the need for customized outreach and policies to enhance voter engagement across diverse U.S. communities."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

@@

The cleaned data will be analyzed and performed using R (@citeR) with `tidyverse` (@citeTidyverse), `here` (@citeHere), `rstanarm` (@citeRstanarm), `modelsummary` (@citeModleSummary), `ggplot2` (@citeGgplot2), `knitr` (@citeKnitr), `marginaleffects` (@citeMarginalEffects), `plotly` (@citePlotly), `tibble` (@citeTibble), `margins` (@citeMargins), `testthat` (@citetestthat) and `kableExtra` (@citeKableExtra). @tbl-clean-data shows a preview of the cleaned dataset.

@@

## Model Set-up

@@

### Model Justification

@@

@@

@@

modelsummary(
  list(
    "Support Biden" = political_preferences1
  ),
  statistic = "mad"
  )
```
Our results are summarized in @tbl-modelresults1. Our results generally matches our expectation. To avoid multicollinearity, the model excluded three variables from each category: race Asian, region Midwest, and employed full-time.The intercept represents the estimated log-odds of supporting Biden when all other predictors are held constant at their reference levels. In this case, the estimated log-odds of supporting Biden for individuals who are Asian, live in the Midwest, and are employed full-time is 0.806.

@@

@fig-modelresults1 (see @sec-credibility-interval) inshows range of coefficient estimates of our model within the 90% probability. Due to the fact that the credibility interval for race "Other" is particularly large, we cannot observe the trend of the 90% credibility intervals of other variables. Therefore we created figure # with the x axis limited from -10 to 10.

Combing @fig-modelresults1 and @fig-modelresults2, we observe statistical significance for the coefficient estimates for students, the Northeast region, individuals identifying with two or more races, individuals identifying with "Other" racial category, Native Americans, Black individuals, and the intercept, Asians in the Midwest region who are employed full-time. The estimates are significant if the intervals do not cross 0. The value for the estimates are in log-odds, indicating that if the coefficient is positive, the individual supports Biden, if negative, the individual supports Trump.

@@

## Racial Variations

@@

## Employment Variations

@@

## Weaknesses and Next Steps

@@

```{r}
#| label: fig-conf
#| fig-cap: Comparing the posterior with the prior 
#| echo: false
#| warning: false
#| message: false
plot(
  political_preferences1,
  "areas"
)
```
```{r}
#| label: fig-conf_lim
#| fig-cap: Comparing the posterior with the prior 
#| echo: false
#| warning: false
#| message: false
#| 
model_plot <- plot(
  political_preferences1,
  "areas"
)
# Modify the x-axis limits
model_plot + xlim(-10, 10) +  # Adjust the limits as needed
  labs(x = "90% Credibility Interval")
```
## Credibility Interval {#sec-credibility-interval}

@@

 # Modify the x-axis limits
 model_plot + xlim(-10, 10) +  # Adjust the limits as needed
   labs(x = "90% Credibility Interval")
```

@@

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

 plot(political_preferences1, "trace")

>>>>

---
title: "Understanding the Socio-Economic Determinants of Voter Behavior: An Analysis of the 2020 U.S. Presidential Election"
author: 
  - Hannah Yu
  - Mary Cheng
  - Yimiao Yuan
thanks: "Code and data are available at: https://github.com/hannahyu07/US-Election"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

The cleaned data will be analyzed and performed using R (@citeR) with `tidyverse` (@citeTidyverse), `here` (@citeHere), `rstanarm` (@citeRstanarm), `modelsummary` (@citeModleSummary), `ggplot2` (@citeGgplot2), `knitr` (@citeKnitr), `marginaleffects` (@citeMarginalEffects), `plotly` (@citePlotly), `tibble` (@citeTibble), `margins` (@citeMargins), and `kableExtra` (@citeKableExtra). @tbl-clean-data shows a preview of the cleaned dataset.

@@

## Model set-up

@@

### Model justification

@@

## Model Implication
In @fig-post_dist, the great fit of the posterior distribution from our logistic regression model with actual voting data suggests accurate capture of voting patterns. This indicates the accuracy of our model's voting preference prediction to the 2020 election data. Also, @fig-post_prior compares the posterior to the prior, which shows minimal parameter changes except for the parameter "raceOther". The discrepancy for the parameter "raceOther" is not critical due to the low representation of "raceOther" in the dataset.

Figure # shows range of coefficient estimates of our model within the 90% probability. Due to the fact that the credibility interval for race "Other" is particularly large, we cannot observe the trend of the 90% credibility intervals of other variables. Therefore we created figure # with the x axis limited from -10 to 10.

Combining figure # and figure #, we observe statistical significance for the coefficient estimates for students, the Northeast region, individuals identifying with two or more races, individuals identifying with "Other" racial category, Native Americans, Black individuals, and the intercept, Asians in the Midwest region who are employed full-time. The estimates are significant if the intervals do not cross 0. The value for the estimates are in log-odds, indicating that if the coefficient is positive, the individual supports Biden, if negative, the individual supports Trump.

The trace plot in @fig-mcmc-1 does not suggest anything out of the ordinary. Also, with the Rhat plot (@fig-mcmc-2), we can observe that everything is close to 1, and no more than 1.05.

@@

Our results are summarized in @tbl-modelresults1.

@@

modelsummary(
  list(
    "Support Biden" = political_preferences1
  ),
  statistic = "mad"
  )
```
Our results generally matches our expectation To avoid multicollinearity, the model excluded three variables from each category: race Asian, region Midwest, and employed full-time.The intercept represents the estimated log-odds of supporting Biden when all other predictors are held constant at their reference levels. In this case, the estimated log-odds of supporting Biden for individuals who are Asian, live in the Midwest, and are employed full-time is 0.806.

@@

@@

## First discussion point

@@

## Third discussion point

@@

## Weaknesses and next steps

@@

```{r}
#| label: fig-conf
#| fig-cap: Comparing the posterior with the prior 
#| echo: false
#| warning: false
#| message: false
plot(
  political_preferences1,
  "areas"
)
```
```{r}
#| label: fig-conf_lim
#| fig-cap: Comparing the posterior with the prior 
#| echo: false
#| warning: false
#| message: false
#| 
model_plot <- plot(
  political_preferences1,
  "areas"
)
# Modify the x-axis limits
model_plot + xlim(-10, 10) +  # Adjust the limits as needed
  labs(x = "90% Credibility Interval")
```

@@

# Modify the x-axis limits
model_plot + xlim(-5, 5) +  # Adjust the limits as needed
  labs(x = "90% Credibility Interval")
```

@@

@fig-trace is a trace plot. It shows if there is existence of signs that the our model runs into issues. We observe lines that are horizontal and oscillating, and have overlaps between the chains. This suggests that there is nothing strange in this trace plot.

@fig-rhat is a Rhat plot. It compares the variability within each chain to the variability between chains in MCMC. Ideally, we want Rhat to be close to 1, which indicates that the chains have converged to the same distribution. We can observe that our Rhat plot are all close to 1, and no more than 1.05. This is a positive sign because it suggests that the MCMC algorithm has reached convergence for our model.

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-trace
#| fig-cap: "Trace plot"

 plot(political_preferences1, "trace")
```

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-rhat
#| fig-cap: "Rhat plot"
